The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-04-11 methods/ms.tex
Try again.

@@

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{jaguirre@nrao.edu}
\newcommand{\lsim}{{_{<}\atop^{\sim}}}
\newcommand{\gsim}{{_{>}\atop^{\sim}}}
\newcommand{\etal}{{et al.\/}}
\newcommand{\ie}{{\em ie.\/}}
\newcommand{\cmq}{cm{$^{-3}$}}
\newcommand{\per}{$^{\rm{-1}}$}
\newcommand{\tc}{{$\theta^1$~Orionis~C}}
%\newcommand{\msol}{M{$_{\odot}$}}
\def\msol{\ifmmode {\>M_\odot}\else {$M_\odot$}\fi}
\newcommand{\lsol}{L{$_{\odot}$}}
\newcommand{\kms}{km~s{$^{-1}$}}
\newcommand{\hii}{H~{\sc ii}}
\newcommand{\Hii}{H~{\sc ii}}
\newcommand{\Ha}{\mbox{H$\alpha$}}
\newcommand{\sii}{S~{\sc ii}}
\newcommand{\Feii}{Fe~{\sc ii}}
\newcommand{\oi}{O~{\sc i}}
\newcommand{\nii}{N~{\sc ii}}
\newcommand{\oiii}{O~{\sc iii}}
\newcommand{\mgii}{Mg~{\sc ii}}
\newcommand{\tco}{{$^{13}$CO}}
\newcommand{\CO}{{$^{12}$CO}}
\newcommand{\Tco}{{$^{12}$CO}}
\newcommand{\co}{C{$^{18}$O}}
\newcommand{\Lsol}{L$_{\odot}$}
\newcommand{\Msol}{M$_{\odot}$}
%\newcommand{\C18o}{C$^{18}$O($1\rightarrow 0$)}
\newcommand{\Check}{{\bf ???}}
\newcommand{\mum}{\ensuremath{\mu \mathrm{m}}}
\newcommand{\flux}{flux density}
\newcommand{\solar}{\ensuremath{\odot}}

@@

%\newpage
%\pagebreak

%\tableofcontents

%\newpage
%\pagebreak

@@

\affil{{$^\penn$}{\it{Department of Physics and Astronomy, University of
      Pennsylvania, Philadelphia, PA }}}
\email{\tt jaguirre@sas.upenn.edu}

@@

\section{Introduction}

@@

The advent of focal plane arrays containing hundreds of individual
bolometers sensitive to millimeter and sub-millimeter (sub-mm)
radiation has enabled large-scale surveys of mm-wavelength continuum
emission from Galactic interstellar dust.  The great advantage of
continuum observations at 1.1mm is that the mass column density does
not depend strongly on the dust temperature $T_d$ (varying only
linearly unless $T_d$ is less than about 10 K) and is independent of
the gas volume density.  Using the same dust opacity as
\citet{enoch06} and \citet{young06}, and assuming a gas-to-dust mass
ratio $X$ of 100, the mass sensitivity can be written as
\begin{equation} 
\label{eq:Mass}
M_{gas}\approx
0.22\left( e^{12.9/T_d}-1 \right) \left({S_\nu\over 15\; {\rm mJy}} \right)
D^{2}_{kpc} \msol 
\end{equation}
where $D_{kpc}$ is the distance to the source in kpc.  (At $S_\nu =
15$~mJy this corresponds to an RMS extinction sensitivity of $A_V\sim
1$ mag, with the same $T_d$ dependence).  It is thus possible with a
wide area, relatively shallow 1.1~mm survey to provide an unmatched,
uniform inventory of massive star-forming and starless condensations.

@@

The outline of the paper is as follows. Section \ref{sec:Observations}
describes the instrument and the observations.  Section
\ref{sec:FluxCalibration} describes the flux calibration and errors.
Section \ref{sec:Astrometry} describes the construction of an absolute
reference frame for the observations and verification against

@@

%\begin{sidewaysfigure}[h]
\begin{figure}
  \begin{minipage}{6.5in}
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
      \caption{$l=-10.5$ to $l=19.5$ Units are Jy/Beam.  The brightest
	sources, e.g. Sgr B2, Sgr A, and sources near $l=10$ and
	$l=13$, appear to be saturated, but this is only a display
	artifact.  The noise is more pronounced from $l=-7$ to $l=-2$
	because this region was observed less.}
    \end{center}
    \label{fig:BGPSMontage}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg2_bw_inv_bar_label} 
      \caption{$l=19.5$ to $l=49.5$.  Units are Jy/Beam .  G34.3+0.15, W
	51, W 43, W 49, and M 17 appear to be saturated, but this is
	only a display artifact}
    \end{center}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg3_bw_inv_bar_label}
      \caption{$l=49.5$ to $l=74.5$.}
    \end{center}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \hspace{-1in}
  \includegraphics[scale=0.8]{fig1_pg5_bw_inv_bar_label} 
  \caption{The Cygnus Arm.}
\end{figure}
%  \vspace{-4in}

@@

\begin{figure}
\hspace{-0.5in}
  \includegraphics[scale=0.8]{fig1_pg4_bw_inv_bar_label} 
  \caption{Cloud complexes centered at $l=111$ in the Perseus Arm.}
\end{figure}

@@

\begin{figure}
  \hspace{-1in}
  \includegraphics[scale=0.8]{fig1_pg6_bw_inv_bar_label} 
  \caption{W3/4/5}
\end{figure}

@@

The observations described here were acquired during six separate
observing sessions at the CSO over the course of two years.  The
observing epochs are given in Table \ref{tab:Observing}, along with
corresponding ranges for the zenith opacity $\tau_{225}$ of the CSO
tipper tau.  Because of the CSO weather multiplexing policy, Bolocam
observations were typically scheduled when $\tau_{225} > 0.06$.
Between each observing epoch, Bolocam was removed from its mount at
the re-imaged Cassegrain focus and stored warm.  Thus the flux
calibration and pointing model must be re-computed for each epoch to
allow for changes in the instrument and optics.

@@

Our basic observing strategy was to raster scan Bolocam by moving the
primary mirror of the CSO to modulate the astrophysical signal faster
than fluctuations in atmospheric opacity.  Each field was scanned with
alternating raster scans along $l$ and $b$.  Starting with Epoch III,
the fundamental scan unit was a $3\arcdeg \times 1\arcdeg$ block.
Each block was scanned with 23 scans along lines of constant $b$, and
67 scans along lines of constant $l$.  In both cases, the spacing
between adjacent rasters was 162\arcsec.  The total time for such a
scan was between 39 and 48 minutes.  In order to fill in the
instanteously undersampled focal plane, we began using in Epoch II a
field rotator to mechanically align the focal plane at an optimal
orientation along the scan direction to improve the sampling
orthogonal to the scan direction.  This is shown in Figure
\ref{fig:Array}.  The data were electronically sampled at 10 Hz along
the scan direction, slightly higher than the Nyquist rate for the scan
speed of 120\arcsec\ s$^{-1}$.  
%CHECK
Without the field rotator, the coverage shows variations of 100\% from
pixel to pixel in a single raster of a field.  With the rotator

The resulting coverage using the
above scan pattern and the field rotator, is typically uniform to
% per Adam, 2009/3/26

 with only a single pass across the field.

@@

\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright
molecular emission lines are shown at their rest frequencies.  Note
that the Bolocam passband rejects $>90\%$ of the $^{12}$CO flux,
leaving our continuum measurements largely uncontaminated.  SO and
CH3OH lines are probably the dominant contributors to line flux in the
passband.  \citet{nummelin1998} found that 22\% of the flux in one
pointing towards Sgr B2 was from lines, and \citet{yoshida2005}
reported $>40\%$ in Orion A from line emission} {fig:Bandpass}{1.0}

\Table
{llll}
{Observing Epochs for the BGPS}
{Begin (UT) & End (UT) & Nights & $\tau_{225}$ }
{tab:Observing}
{
2005 Jul 03 & 2005 Jul 13 & 5 & 0.1 \\
2005 Sep 04 & 2005 Sep 12 & 5 & 0.1 \\
2006 Jul 01 & 2006 Jul 13 & 5 & 0.1 \\
2006 Sep 01 & 2006 Sep 09 & 5 & 0.1 \\
2007 Jul 01 & 2007 Jul 13 & 5 & 0.1 \\
2007 Sep 01 & 2007 Sep 09 & 5 & 0.1 \\
}

@@

Comparison of the accuracy of the Bolocam flux density calibration
with the SCUBA 850 \mum\ results is complicated by the unknown effect
of the dust spectral index.  However, there are several large-area
surveys with the MAMBO instrument at 1.2 mm with which we can compare.
Figure \ref{fig:CygnusComparison} shows the comparison of fluxes
extracted from our map in the Cygnus Arm with those of
\citet{motte07}.

@@

Relative alignment was performed by finding the peak of the
cross-correlation between images and a pointing master selected from
the epoch with the best-constrained pointing model for that
field. Each observation was initially mapped individually, then all
observations of a given field were cross-correlated with a selected
master image of that field. The cross-correlation peak was fit with a
gaussian and the difference between the gaussian peak and the image
center was used as the pixel offset. The offsets were recorded and
written to the timestreams. Finally, all observations of a field were
merged into a single timestream with pointing offsets applied to
create the field mosaic.

This method of alignment avoids the ambiguities inherent in using
extracted sources to align fields, as the Bolocam sources are rarely
point-like.  It further avoids the pointing smearing and loss of peak
flux density which would result if the the maps were coadded using the
pointing model alone.

@@

\subsection{Comparison to SCUBA Legacy Source Positions}
\label{sec:SCUBAPointingComparison}
%and SHARC-II Positions}

The same cross-correlation procedure was used to compare to the SCUBA
maps from the Legacy Catalog.  The morphological correlation between
SCUBA and BGPS sources is excellent, so cross-correlation maps could
be used to test our pointing model.

An example is shown in Figure \ref{fig:SCUBAPointingComparison}.

{\it Do the negative bowls of either survey bunge up the
cross-correlation?  If not, how do we deal with them?}

For the purposes of comparison of both flux and pointing with other
surveys, we present in Appendix \ref{app:ReferenceSources} a selection
of bright, compact, relatively isolated sources.

@@

% 8
% -----------------------------------------------------------------------------
\FigureTwo{plane_astrometry}{pointing_model_skydist}{Left: The
distribution of absolute reference sources (crosses) in the Galactic
Plane (Appendix \ref{app:PointingCalibrators}).  Also shown (crosses)
are the locations of bright, isolated, compact sources for comparison
with other surveys, as given in Appendix \ref{app:ReferenceSources}.
Right: The distribution of the pointing calibration sources across the
local sky in Hawaii during the July 2007 epoch when the pointing model
was constructed.  Notice the good sampling of the entire region used
for the Survey.}{fig:PointingCalibrators}{1.0}
% -----------------------------------------------------------------------------

@@

  \subfigure[{\it Left}: the Bolocam map in a portion of the $l=30$
  field.  {\it Right}: the SLC maps in the same region.  The green
  contours on the Bolocam map indicate the positions of the SLC
  sources.]
  {\label{fig:SCUBAPointingComparison-a}\includegraphics[scale=0.8]
  {scuba_bolocam_l030}}

@@

We assume the raw timestream data $d$ for each bolometer (indexed by
$i$) can be written as
\begin{equation}
d_i(t) = s_i(t)+a_i(t)+e_i(t)+\epsi_i(t)
\end{equation}
where $s$ is the astrophysical signal, $a$ is atmospheric fluctuation
noise, $e$ are non-random signals due to the instrument itself, and
$\epsi$ is irreducible Gaussian noise due to photon shot and detector
noise.  The process of making a maximum likelihood map in the presence
of Gaussian noise is well-understood, and if the only contributions to
the data were $s$ and $\epsi$, then the minimum variance estimator of
the true astrophysical map is given by
\begin{equation}
\label{eq:LSMap}
m = (A^T W A) A^T W d
\end{equation}
where $A$ encodes the pointing information and $W$ is the covariance
matrix of the noise $\epsi$.  Note that the mapping from data to map
is a linear operator, $M =(A^T W A) A^T W$.  For compactness, Equation
\ref{eq:LSMap} will be written as $M[d] = m$.  Note that the mapping
operation is not invertible.  However, note that given a map $m$,
there exists a linear operation which makes predictions about the
observed timestreams, namely $d = A m$; this will be denoted $T[m] =
d$.

@@

We proceed iteratively as follows. 
\begin{equation}
\label{eq:FirstIterStep}
\tilde{a}^{(n)}(t) = 
\sum_i{d_i(t)-\tilde{s}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)}
\end{equation}
where $\tilde{s}^{(0)}_i(t)=0$.  This mean atmosphere model is then
fit to each bolometer to obtain the relative gains (flat field) of the
detectors
\begin{equation}
\tilde{a}^{(n)}_i(t) = \frac{d_i(t) \cdot a(t)}{d_i(t)^2} a(t) = r_i a(t)
\end{equation}
This is illustrated in Figure \ref{fig:Flatfield}.  The instrument
errors, which are typically smaller than the atmosphere, are then
estimated
\begin{equation}
\tilde{e}^{(n)}_i(t) = some models
\end{equation}
and both the atmosphere and instrument models are subtracted from the
raw time series
\begin{equation}
d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)
\approx s_i(t) + \epsi_i(t)
\end{equation}
This is the best estimate of signal plus irreducible noise at
iteration step $n$, and is made into a map
\begin{equation}
M[d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)] = \tilde{m}^{(n)}
\end{equation}
The current best map $\tilde{m}^{(n)}$ is then deconvolved to provide
a relatively low-noise, smooth map from which to generate a timestream
\begin{equation}
T[{\cal D}[\tilde{m}^{(n)}]] = \tilde{s}^{(n)}_i(t)
\end{equation}
We note that it is possible to produce an ``error map''
\begin{equation}
N = M[d_i(t) - \tilde{s}^{(n)}_i(t) - \tilde{a}^{(n)}_i(t) - 
\tilde{e}^{(n)}_i(t)]
\end{equation}
The process then begins again with Equation \ref{eq:FirstIterStep}.

@@

@@

Fortunately, most of the AC powerline pick-up occurs at frequencies
where there is no astrophysical signal, given the beam size and scan
speed.  Thus this error is dealt with by first notch filtering at the
line frequencies and then low-pass filtering the data.  Because of the
low correlation with astrophysical signal, this step is performed only
once and is not iterated.

@@

%\clearpage 
%
%\Figure{deconvolution_comparison}{The fractional flux lost as a function of
%source size for well-separated Gaussian sources with the FWHM as
%indicated.  Legend: Size of deconvolution kernel used to make model }{fig:deconv_filter}{1.0}

\section{Final Maps}

@@

\section{Summary}

>>>>

% Major Sections
% sec:Introduction
% sec:Observations
% sec:FluxCalibration
% sec:Astrometry
% sec:Mapping

@@

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{{\tt jaguirre@sas.upenn.edu}}
\newcommand{\lsim}{{_{<}\atop^{\sim}}}
\newcommand{\gsim}{{_{>}\atop^{\sim}}}
\newcommand{\etal}{{et al.\/}}
\newcommand{\ie}{{\em ie.\/}}
\newcommand{\cmq}{cm{$^{-3}$}}
\newcommand{\per}{$^{\rm{-1}}$}
\newcommand{\tc}{{$\theta^1$~Orionis~C}}
%\newcommand{\msol}{M{$_{\odot}$}}
\def\msol{\ifmmode {\>M_\odot}\else {$M_\odot$}\fi}
\newcommand{\lsol}{L{$_{\odot}$}}
\newcommand{\kms}{km~s{$^{-1}$}}
\newcommand{\hii}{H~{\sc ii}}
\newcommand{\Hii}{H~{\sc ii}}
\newcommand{\Ha}{\mbox{H$\alpha$}}
\newcommand{\sii}{S~{\sc ii}}
\newcommand{\Feii}{Fe~{\sc ii}}
\newcommand{\oi}{O~{\sc i}}
\newcommand{\nii}{N~{\sc ii}}
\newcommand{\oiii}{O~{\sc iii}}
\newcommand{\mgii}{Mg~{\sc ii}}
\newcommand{\tco}{{$^{13}$CO}}
\newcommand{\CO}{{$^{12}$CO}}
\newcommand{\Tco}{{$^{12}$CO}}
\newcommand{\co}{C{$^{18}$O}}
\newcommand{\Lsol}{L$_{\odot}$}
\newcommand{\Msol}{M$_{\odot}$}
%\newcommand{\C18o}{C$^{18}$O($1\rightarrow 0$)}
\newcommand{\Check}{{\bf ???}}
\newcommand{\mum}{\ensuremath{\mu \mathrm{m}}}
\newcommand{\flux}{flux density}
\newcommand{\solar}{\ensuremath{\odot}}

@@

\newpage
\pagebreak

\tableofcontents

\newpage
\pagebreak

@@

\affil{{$^\penn$}{\it{Department of Physics and Astronomy, University of
      Pennsylvania, Philadelphia, PA }}}
\email{\myemail}

@@

\section{Introduction}
\label{sec:Introduction}

@@

The advent of focal plane arrays containing hundreds of individual
bolometers sensitive to millimeter and sub-millimeter (sub-mm)
radiation has enabled large-scale surveys of mm-wavelength continuum
emission from Galactic interstellar dust.  The great advantage of
continuum observations at 1.1mm is the low optical depth of the dust,
permitting an estimate of the mass of the emitting region.  Further,
the mass column density is independent of the gas volume density.

Using the same dust opacity as \TBD\ ({\it put in original references})
%\citet{enoch06} and \citet{young06},
and assuming a gas-to-dust mass ratio $X$ of 100, the mass sensitivity
can be written as
\begin{equation} 
\label{eq:Mass}
M_{gas}\approx
0.22\left( e^{12.9/T_d}-1 \right) \left({S_\nu\over 15\; {\rm mJy}} \right)
D^{2}_{kpc} \msol 
\end{equation}
where $D_{kpc}$ is the distance to the source in kpc.  (At $S_\nu =
15$~mJy this corresponds to an RMS extinction sensitivity of $A_V\sim
1$ mag, with the same $T_d$ dependence).  For Bolocam, $h\nu/k \approx
13$~K, so for $T_d > XXX$ the dependence is approximately linear.  At
the coldest temperatures expected, the mass changes a factor of 2.9
between 20 and 10 K.  Far more important is the accurate determination
of the distance to the clump.

Thus it is thus possible with a wide area, relatively shallow 1.1~mm
survey to provide an unmatched, uniform inventory of massive
star-forming and starless condensations.

@@

%\section{Overview of the Analysis}
%\label{sec:Overview}

{\it material from the Letter here}

A number of different descriptions of the reduction of Bolocam data
have been published \citep{laurent05,enoch06,sayers09}.  This work
addresses the issues specific to the BGPS, significantly improving the
method of \citet{enoch06} for generating maps and characterizing their
properties.  It incorporates also some of the technical developments
of \citet{sayers09}, though the exquisite control of systematics and
noise of that analysis is neither necessary nor possible here.

The outline of the paper is as follows. Section \ref{sec:Observations}
describes the instrument and the observations.  Section
\ref{sec:FluxCalibration} describes the flux calibration and errors.
Section \ref{sec:Astrometry} describes the construction of an absolute
reference frame for the observations and verification against other
(sub)millimeter catalogs.  Section \ref{sec:Mapping}, \ref{sec:Noise},
\ref{sec:Photometry}, \ref{sec:FinalMaps}, \ref{sec:Discussion}.

@@

%\begin{sidewaysfigure}[h]
\begin{figure}
  \begin{minipage}{6.5in}
    \begin{center}
 %     \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
      \caption{$l=-10.5$ to $l=19.5$ Units are Jy/Beam.  The brightest
	sources, e.g. Sgr B2, Sgr A, and sources near $l=10$ and
	$l=13$, appear to be saturated, but this is only a display
	artifact.  The noise is more pronounced from $l=-7$ to $l=-2$
	because this region was observed less.}
    \end{center}
    \label{fig:BGPSMontage}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
 %     \includegraphics[scale=0.9]{fig1_pg2_bw_inv_bar_label} 
      \caption{$l=19.5$ to $l=49.5$.  Units are Jy/Beam .  G34.3+0.15, W
	51, W 43, W 49, and M 17 appear to be saturated, but this is
	only a display artifact}
    \end{center}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
%      \includegraphics[scale=0.9]{fig1_pg3_bw_inv_bar_label}
      \caption{$l=49.5$ to $l=74.5$.  \TBD\ Add annotations}
    \end{center}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \hspace{-1in}
%  \includegraphics[scale=0.8]{fig1_pg5_bw_inv_bar_label} 
  \caption{The Cygnus Arm. \TBD\ Add annotations}
\end{figure}
%  \vspace{-4in}

@@

\begin{figure}
\hspace{-0.5in}
%  \includegraphics[scale=0.8]{fig1_pg4_bw_inv_bar_label} 
  \caption{Cloud complexes centered at $l=111$ in the Perseus Arm. \TBD\ Add annotations}
\end{figure}

@@

\begin{figure}
  \hspace{-1in}
%  \includegraphics[scale=0.8]{fig1_pg6_bw_inv_bar_label} 
  \caption{W3/4/5 \TBD\ Add annotations}
\end{figure}

@@

The observations described here were acquired during six separate
observing sessions at the CSO over the course of two years.  The
observing epochs are given in Table \ref{tab:Observing}, along with
corresponding ranges for the zenith opacity $\tau_{225}$ of the CSO
tipper tau.  Because of the CSO weather multiplexing policy, Bolocam
observations were typically scheduled when $\tau_{225} > 0.06$.
Between each observing epoch, Bolocam was removed from its mount at
the re-imaged Cassegrain focus and stored warm.  Thus the flux
calibration and pointing model must be re-computed for each epoch to
allow for variations in the instrument and optics.  We found that the
flux calibration did not in fact differ significantly between epochs.
For the pointing model, we computed a well-constrained version for
Epoch V and aligned all images to this.

@@

Our basic observing strategy was to raster scan Bolocam by moving the
primary mirror of the CSO to modulate the astrophysical signal faster
than fluctuations in atmospheric opacity.  Each field was scanned with
alternating raster scans along $l$ and $b$.  Starting with Epoch III,
the fundamental scan unit was a $3\arcdeg \times 1\arcdeg$ block.
Each block was scanned with 23 scans along lines of constant $b$, and
67 scans along lines of constant $l$.  In both cases, the spacing
between adjacent rasters was 162\arcsec.  The total time for such a
scan was between 39 and 48 minutes.  The data were electronically
sampled at 10 Hz along the scan direction, slightly higher than the
Nyquist rate for the scan speed of 120\arcsec\ s$^{-1}$.

In order to fill in the instanteously undersampled focal plane, we
began using in Epoch II a field rotator to mechanically align the
focal plane at an optimal orientation along the scan direction to
improve the sampling orthogonal to the scan direction.  This is shown
in Figure \ref{fig:Array}.  Achieving good sampling is complicated by
a number of missing pixels.  A simulation was performed to determine
the optimal angle to rotate the array with respect to the scan
direction to account for the beam spacing and missing pixels.  During
the turnaround following a raster, the field rotator is adjusted to
this optimum angle.
%CHECK
% per Adam, 2009/3/26
Without the field rotator, the coverage shows variations of 100\% from
pixel to pixel in a single raster of a field.  With the rotator, this
is reduced to $\sim40\%$.

@@

\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright
molecular emission lines are shown at their rest frequencies.  Note
that the Bolocam passband rejects $>90\%$ of the $^{12}$CO flux,
leaving our continuum measurements largely uncontaminated.  SO and
CH$_3$OH lines are probably the dominant contributors to line flux in
the passband.  \citet{nummelin1998} found that 22\% of the flux in one
pointing towards Sgr B2 was from lines, and \citet{yoshida2005}
reported $>40\%$ in Orion A from line emission. \TBD\ Other lines
lying in the Bolocam passband include CS($5\to4$) and ($6\to5$) (245
and 293 GHz), HCN($3\to2$) (265 GHz) and HCO$^+(3\to2)$ (267 GHz).}
{fig:Bandpass}{1.0}

\Table
{lllll}
{Observing Epochs for the BGPS}
{Number & Begin (UT) & End (UT) & Nights & $\tau_{225}$ }
{tab:Observing}
{
I   & 2005 Jul 03 & 2005 Jul 13 & 5 & 0.1 \\
II  & 2005 Sep 04 & 2005 Sep 12 & 5 & 0.1 \\
III & 2006 Jul 01 & 2006 Jul 13 & 5 & 0.1 \\
IV  & 2006 Sep 01 & 2006 Sep 09 & 5 & 0.1 \\
V   & 2007 Jul 01 & 2007 Jul 13 & 5 & 0.1 \\
VI  & 2007 Sep 01 & 2007 Sep 09 & 5 & 0.1 \\
}

@@

Comparison of the accuracy of the Bolocam flux density calibration
with the SCUBA 850 \mum\ results is complicated by the unknown effect
of the dust spectral index.  However, there are several large-area
surveys with the MAMBO instrument at 1.2 mm with which we can compare.
Figure \ref{fig:CygnusComparison} shows the comparison of fluxes
extracted from our map in the Cygnus Arm with those of
\citet{motte07} (M07).

The figures of the Cygnus-X region presented by M07
%Motte et al. (2009)
, obtained at a central wavelength of 1.2 mm using the 30 meter IRAM
telescope, appear nearly identical to the 1.1 mm Bolocam images.  The
BGPS mapping strategy, larger beam, and field of view provides better
recovery of the large scale structure of the dust continuum emission,
but inferior angular resolution.  Nevertheless, there is excellent
agreement between the structure revealed by the two data sets.

The Bolocam and M07 data sets have flux calibration consistent within
about 20\%.  However, direct comparison with the fluxes listed in M07
is difficult due to the factor of three difference in the size of the
beams used.  For example, the the brightest source in the Cyg-X region
is DR 21.  While M07 quote a peak flux density per beam of
4.2 Jy/beam for the brightest feature in the DR 21 complex (their peak
N46) and 17.9 Jy/beam in a 0.19 x 0.14 pc box, BGPS measures a peak
flux of 15.5 Jy/beam (the beam being roughly 31\arcsec\ in diameter).
Fluxes are best compared using sources compact as observed with the
beam of the IRAM 30 meter telescope.

We have convolved the fits files\footnote{{\tt
ftp://cdsarc.u-strasbg.fr/pub/cats/J/A+A/476/1243/fits/}} with a
Gaussian kernel that generates an approximately 35\arcsec\ beam,
matched to our images (mapped without application of the distortion
correction).  The results are shown in Figure
\ref{fig:CygnusComparison}.

\Table
{lll}
{Flux density comparison with \citet{motte07}.}
{
Source  &   BGPS           &  Motte \\
        &   peak mJy/beam  &  mJy in box (see Motte+ Table 1)
}
{tab:CygnusComparison}
{
 N11    &   350 - 550      &  340 \\
 N12    &   1660-2200      &  1630 \\
 N16    &   613            &  550 \\
 N18    &   618            &  430 \\
 N26    &   314            &  340 \\
 N46    &   15500          &  17930  (DR21 S)
}

@@

Relative alignment between different observations of the same field
was performed by finding the peak of the cross-correlation between
images and a pointing master selected from the epoch with the
best-constrained pointing model for that field. (In almost all cases,
this was from Epoch V.)  Each observation was initially mapped
individually, then all observations of a given field were
cross-correlated with a selected master image of that field. The
cross-correlation peak was fit with a gaussian and the difference
between the gaussian peak and the image center was used as the pixel
offset. 
%The offsets were recorded and written to the
%timestreams. 
Finally, all observations of a field were merged into a single
timestream with pointing offsets applied to create the field mosaic.

This method of alignment avoids the ambiguities inherent in using
extracted sources to align fields, as the Bolocam sources are rarely
point-like.  It further avoids the pointing smearing and loss of peak
flux density which would result if the the maps were coadded using the
pointing model alone, where each individual observation would be
coadded with tthe 6\arcsec\ RMS uncertainty of the model.  The
relative alignment using the correlation maps is typically accurate to
\TBD\ $<1\arcsec$.

@@

\subsection{Comparison to SCUBA Legacy Catalog Source Positions}
\label{sec:SCUBAPointingComparison}
%and SHARC-II Positions}

All of the SCUBA 850 and 450 \mum\ data has been re-processed in a
uniform manner and made publicly available as the SCUBA Legacy Catalog
(hereafter SLC).  These maps provide an excellent cross-check on the
accuracy of the BGPS pointing model.  We have used the same
cross-correlation procedure as that used to obtain the relative
alignment of Bolocam observations to compare to the SLC.  A number of
concerns might arise in comparing the Bolocam to the SCUBA maps,
including the differing scan strategies (SCUBA maps were typically
taken in jiggle-map mode, which tends to remove extended structure)
and the often restricted range of positions measured within an
extended cloud for the SLC.  The cross-correlation technique actually
mitigates these effects, since it does not require the establishment
of a single position for an object (which may differ due to the change
in wavlength between the surveys or due to the observing strategy) but
rather makes use of all the information available.  In the event, the
morphological correlation between SCUBA and BGPS sources is generally
excellent, giving a high degree of confidence to the comparison and to
the results obtained from the pointing model.  An example is shown in
Figure \ref{fig:SCUBAPointingComparison}.  We find that the RMS offset
between BGPS and SLC positions within a $3\arcdeg \times 1\arcdeg$
bloc defined by the BGPS is \TBD\ a few arcseconds, consistent with
the errors derived for the Bolocam pointing model.

For the purposes of future comparison of both flux and pointing with
other surveys, we present in Appendix \ref{app:ReferenceSources} a
selection of bright, compact, relatively isolated sources.

@@

% 8
% -----------------------------------------------------------------------------
\FigureTwo{plane_astrometry}{pointing_model_skydist}{Left: The
distribution of absolute reference sources (crosses) in the Galactic
Plane (Appendix \ref{app:PointingCalibrators}).  Also shown (crosses)
are the locations of bright, isolated, compact sources for comparison
with other surveys, as given in Appendix \ref{app:ReferenceSources}.
Right: The distribution of the pointing calibration sources across the
local sky in Hawaii during the July 2007 epoch when the pointing model
was constructed.  Note the good sampling built up over the range of
sky observed in the Survey.}{fig:PointingCalibrators}{1.0}
% -----------------------------------------------------------------------------

@@

  \subfigure[{\it Left}: the Bolocam map in a portion of the $l=30$
  field.  {\it Right}: the SLC maps in the same region.  The green
  contours on the Bolocam map indicate the positions of the SLC
  sources.  \TBD\ {\it Do the negative bowls of either survey bunge up
  the cross-correlation?  If not, how do we deal with them?  The SCUBA
  image here has been somewhat doctored because of a large DC offset
  in the diagonal filament at center}]
  {\label{fig:SCUBAPointingComparison-a}\includegraphics[scale=0.8]
  {scuba_bolocam_l030}}

@@

We assume the raw timestream data $d$ for each bolometer (indexed by
$i$) can be written as
\begin{equation}
d_i(t) = s_i(t)+a_i(t)+e_i(t)+\epsi_i(t)
\end{equation}
where $s$ is the astrophysical signal, $a$ is atmospheric fluctuation
noise, $e$ are non-random signals due to the instrument itself, and
$\epsi$ is irreducible Gaussian noise due to photon shot and detector
noise.  The process of making a maximum likelihood map in the presence
of Gaussian noise is well-understood, and if the only contributions to
the data were $s$ and $\epsi$, then the minimum variance estimator of
the true astrophysical map is given by
\begin{equation}
\label{eq:LSMap}
m = (A^T W A) A^T W d
\end{equation}
where $A$ encodes the pointing information and $W$ is the inverse of
the covariance matrix of the time domain noise $\epsi$,
\begin{equation}
W = {\langle \epsi_i \epsi_j \rangle}^{-1}
\end{equation}
In this case, the mapping from data to map is a linear operator, $M
=(A^T W A) A^T W$.  For compactness, Equation \ref{eq:LSMap} will be
written as $M[d] = m$.  Note that the mapping operation is not
invertible.  However, given a map $m$ and an observing matrix $A$,
there exists a linear operation which makes predictions about the
observed timestreams, namely $d = A m$; this will be denoted $T[m] =
d$.

@@

We proceed iteratively, estimating $a$ and $e$ in the order of their
relative strengths, beginning with the atmosphere.  The first
iteration step is then
\begin{equation}
\label{eq:FirstIterStep}
\tilde{a}^{(n)}(t) = 
\sum_i{d_i(t)-\tilde{s}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)}
\end{equation}
where $\tilde{s}^{(0)}_i(t)=0$.  This mean atmosphere model is then
fit to each bolometer to obtain the relative gains (flat field) of the
detectors
\begin{equation}
\tilde{a}^{(n)}_i(t) = \frac{d_i(t) \cdot a(t)}{d_i(t)^2} a(t) = r_i a(t)
\end{equation}
This is illustrated in Figure \ref{fig:Flatfield}.  The instrument
errors, which are typically smaller than the atmosphere, are then
estimated
\begin{equation}
\tilde{e}^{(n)}_i(t) = some models
\end{equation}
and both the atmosphere and instrument models are subtracted from the
raw time series
\begin{equation}
d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)
\approx s_i(t) + \epsi_i(t)
\end{equation}
This is the best estimate of signal plus irreducible noise at
iteration step $n$, and is made into a map
\begin{equation}
M[d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)] = \tilde{m}^{(n)}
\end{equation}
The current best map $\tilde{m}^{(n)}$ is then deconvolved to provide
a relatively low-noise, smooth map from which to generate a timestream
\begin{equation}
T[{\cal D}[\tilde{m}^{(n)}]] = \tilde{s}^{(n)}_i(t)
\end{equation}
where ${\cal D}$ represents the deconvolution operation.  At this
stage, the iterative process could begin again with Equation
\ref{eq:FirstIterStep}, though it is highly useful to produce an
``error map''
\begin{equation}
E = M[d_i(t) - \tilde{s}^{(n)}_i(t) - \tilde{a}^{(n)}_i(t) - 
\tilde{e}^{(n)}_i(t)]
\end{equation}
This error map serves as a straightforward visualization of the
progress of the iterative method, since in a perfect process it would
be map of $\epsi$.

@@

To further remove the atmosphere (and other correlated instrument
noise) without a detailed physical model, we use a principal component
analysis (PCA) as given in \citet{laurent05}.

@@

Fortunately, most of the AC powerline pick-up occurs at frequencies
where there is no astrophysical signal, given the beam size and scan
speed.  Thus this error is dealt with by first notch filtering at the
line frequencies and then low-pass filtering the data.  Because of the
low correlation with astrophysical signal, this step is performed only
once and is not iterated (c.f. Figure XXX of \citet{sayers09}).

@@

\section{Final Maps and Data Release}
\label{sec:FinalMaps}

@@

\section{Discussion}
\label{sec:Discussion}
